---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

This (very small) package provides a function to reapportionate data from one set of spatial polygons to another. This is useful, e.g., when one has some data which is distributed in some administrative units (e.g., the IRIS, the infra-urban unit at which the INSEE, the French national statistics institute, releases the census data) and needs to have it in some other units (e.g., "bureaux de vote", the French polling stations). 

At the moment, the package only offers the option to reapportion data by hypothesising homogeneous repartition of the variable inside a unit. I have plans to implement other schemes in the future. If you are aware of some alternative schemes, please [submit an issue](https://github.com/joelgombin/spReapportion/issues). 

# Installation

This package is not on CRAN. For the time being, install it via the `devtools` package:

```{r install, eval=FALSE}
devtools::install_github("joelgombin/spReapportion")
```

# Usage

There is only one function for the time being in this package. The usage is as such:

```{r usage}
library(spReapportion)
data(ParisPollingStations2012)
data(ParisIris)
data(RP_2011_CS8_Paris)
CS_ParisPollingStations <- spReapportion(ParisIris, ParisPollingStations2012, RP_2011_CS8_Paris, "DCOMIRIS", "ID", "IRIS")

```

You can then make sure that everything went fine, for example by comparing maps of the same data before and after reapportionment. Here, I'll use the great `tmap` package.

```{r mapping}
library(tmap)
library(dplyr)

# first transform counts to percentages

RP_2011_CS8_Paris <- RP_2011_CS8_Paris %>%
  mutate_each(funs(. / C11_POP15P * 100), C11_POP15P_CS1:C11_POP15P_CS8)

# add data to the SPDF
ParisIris <- append_data(ParisIris, RP_2011_CS8_Paris, key.shp = "DCOMIRIS", key.data = "IRIS")

# plot original census units (IRIS)

tm_shape(ParisIris) +
  tm_fill(col = "C11_POP15P_CS3", palette = "Blues", style = "quantile", n = 6, title = "Independent workers") +
  tm_borders()

```

```{r mapping2}
# transform counts to percentages in the reapportionned data
CS_ParisPollingStations <- CS_ParisPollingStations %>%
  mutate_each(funs(. / C11_POP15P * 100), C11_POP15P_CS1:C11_POP15P_CS8)

ParisPollingStations2012 <- append_data(ParisPollingStations2012, CS_ParisPollingStations, key.shp = "ID", key.data = "new_ID")

tm_shape(ParisPollingStations2012) +
  tm_fill(col = "C11_POP15P_CS3", palette = "Blues", style = "quantile", n = 6, title = "Independent workers") +
  tm_shape(ParisIris) +
  tm_borders()


```

